---
title: "BC School Tracker"
author: "Jens von Bergmann"
date: "Last updated at `r format(Sys.time(), '%d %B, %Y - %H:%M',tz='America/Vancouver')`"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 5,
	fig.retina = 2,
	fig.width = 7,
	message = FALSE,
	warning = FALSE,
	dpi = 150
)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(rvest)
library(mountainmathHelpers)
library(sf)
library(PROJ)
library(cancensus)
library(lubridate)
library(sanzo)
source(here::here("R/helpers.R"))

options("cancensus.cache_path"=here::here("data"))
ha_colours <- setNames(c(trios$c157,trios$c149),
                       c("Fraser","Rest of BC","Vancouver Coastal" , "Vancouver Island", "Interior", "Northern"))

extract_table_data <- function(node){
    h <- node %>% rvest::html_nodes("thead th") %>% rvest::html_text()
  rows <- node %>% rvest::html_nodes("tbody tr")
  data <- rows %>% lapply(function(d) d %>% 
                            rvest::html_nodes("td") %>% 
                            rvest::html_text() %>% 
                            t() %>% 
                            as.data.frame) %>%
    bind_rows() %>%
    setNames(h)
  
  data
}
```

```{r}
get_data_for_page <- function(p=1){
  url = "https://us-east-1-renderer-read.knack.com/v1/scenes/scene_1/views/view_3/records?callback=jQuery17201415845315599671_1606757085608&format=both&rows_per_page=100&sort_field=field_16&sort_order=desc&_=1606757085628"
  url <- paste0(url,"&page=",p)
  
  d <- httr::GET(url,httr::set_cookies( "connect.sid"="s%3A51R3S-8YTv08QV_IejVfqJsW1RxnUdln.WuHEEIrAF6niDEAB3MWgjvWA%2FkzArewbBDZ%2FppCUdVY"),
                 httr::accept("text/javascript, application/javascript, application/ecmascript, application/x-ecmascript, */*; q=0.01"),
                 httr::add_headers("X-Knack-Application-Id"= "5faae3b10442ac00165da195",
                                   "Accept-Encoding"="gzip, deflate, br",
                                   "X-Knack-REST-API-Key"= "renderer",
                                   "x-knack-new-builder"= "true"))
  
  c <- httr::content(d,"text") %>%
    gsub("^.*\\(\\{","{",.) %>%
    gsub("\\}\\);$","}",.) %>%
    jsonlite::fromJSON()
  
  data <- c$records %>% as_tibble()
  
  #print(paste0("Got data for page ",c$current_page," of ",c$total_pages))

  
  attr(data,"total_pages") <- c$total_pages
  data 
}

clean_data <- function(d){
  tibble(Date=as.Date(d$field_16,format="%m/%d/%Y"),
           Name=d$field_13_raw %>% lapply(function(f)f$identifier) %>% as.character,
           `Health Authority`=d$field_26_raw %>% lapply(function(f)f$identifier) %>% as.character,
           Verification=d$field_14_raw %>% lapply(function(f)f$identifier) %>% as.character,
           `Exposure dates`=d$field_15,
           `Exposure count`=d$field_25,
         #Status =d$field_30,
         Variant=d$field_38) %>%
      bind_cols(d$field_19_raw %>% as_tibble) %>%
    mutate(E=str_extract(`Exposure count`,"Exposure \\d+|Secondary \\d+") %>% 
             unlist() %>% 
             gsub("Exposure |Secondary ","",.) %>% 
             as.integer()) %>%
    mutate_at(c("latitude","longitude"),as.numeric)
}

get_all_data <- function(){
  p=1
  raw_data <- get_data_for_page(p)
  data <- clean_data(raw_data)
  while (attr(raw_data,"total_pages")>p) {
    p <- p + 1
    raw_data <- get_data_for_page(p)
    data <- bind_rows(data, clean_data(raw_data))
  }
  data
}

extract_first_exposure_date <- function(ed){
  ed %>% 
    gsub("<p>|</p>|<br />|Exposure date |Exposure dates |Exposures dates |<hr />","",.)  %>% 
    strsplit(",|and") %>% 
    map(trimws) %>% 
    map(function(d)d[d!=""]) %>% 
    map(first) %>% 
    gsub("-.+| \\(.+| \\[.+","",.) %>%
    unlist
}

add_first_exposure_date <- function(data){
  data %>%
    mutate(fe=extract_first_exposure_date(.data$`Exposure dates`))
}

data <- get_all_data()
```

In BC we don't have good data on community level COVID cases. But we can, somewhat cynically, use school exposures as a proxy. The [BC School Covid Tracker project](https://bcschoolcovidtracker.knack.com/bc-school-covid-tracker#home/) has been keeping count of school exposures and meticulously documenting and geocoding the data. That makes is easy for use to create a heatmap of school exposures.

The code for this notebook is [available for anyone to adapt and use for their own purposes](https://github.com/mountainMath/BCCovidSnippets/blob/main/bc_school_tracker.Rmd).

## Variants of concern
The share of school exposures that are related to variants of concern can give an indication of the general proliferation of variants of concern.

We look at the share of school exposures involving variants of concern for each week, where we split the week between Monday and Tuesday as exposure notifications coming out on Monday usually relate to exposures in the preceding week. Variant of concern screening takes some extra time and exposures might get identified as realting to variants of concern a couple of days later, so shares in the most recent week (and sometimes also the week before that) may rise as exposures gets updated. In particular, the lag will bias the share of exposures involving variants of concern in the current week downward.

Shares are computed separately for each Health Authority, we only show Health Authorities that have flagged exposure events as relating to variants of concern.

```{r schools_voc}
plot_data <- data %>%
  mutate(`Health Authority`=recode(`Health Authority`,
                                   "Fraser Health Authority"="Fraser",
                                   "Interior Health Authority"="Interior",
                                   "Vancouver Coastal Health"="Vancouver Coastal",
                                    "Vancouver Island Health Authority"="Vancouver Island",
                                   "Northern Health"="Northern")) %>%
  mutate(VOC=grepl("concern|variant",`Variant`,ignore.case = TRUE)) %>%
  mutate(Week=ceiling_date(Date,"week",week_start=2)-1) %>%
  count(Week,VOC,`Health Authority`) %>%
  complete(Week,VOC,`Health Authority`,fill=list(n=0)) %>%
  group_by(Week,`Health Authority`) %>%
  mutate(share=n/sum(n)) %>%
  #filter(Week >= min(filter(.,VOC)$Week)) %>%
  filter(VOC) %>%
  filter(Week >= min(filter(.,VOC,n>0)$Week)) %>%
  ungroup %>%
  mutate(week=factor(strftime(Week,"%b %d"),levels=sort(unique(Week)) %>% strftime(.,"%b %d")),
         w=strftime(Week-2,format = "%U") %>% as.integer) 

plot_data %>%
  group_by(`Health Authority`) %>%
  filter(sum(n)>0) %>%
ggplot(aes(x=week,y=share,fill=`Health Authority`)) +
  geom_bar(stat="identity") +
  scale_y_continuous(labels=scales::percent) +
  scale_fill_manual(values=ha_colours,guide=FALSE) +
  facet_wrap("`Health Authority`",ncol=2) +
  labs(title="School exposures involving variants of concern",
       x="Initial exposure letter in week ending",
       y="Share of exposures involving variant of concern",
       caption="MountainMath, Data: BC School Covid Tracker")
```



## School Exposure Heat Map
This data is not normalized by population, so it should be viewd in context of the school aged (5-17yo) population further down.

```{r school-tracker-schools}
bb<-metro_van_bbox()
m <- cancensus::get_census("CA16",regions=list(CMA="59933"),geo_format = "sf",level="CSD",quiet = TRUE)


school_data <- data %>%
  mutate_at(c("latitude","longitude"),as.numeric) %>%
  filter(!is.na(latitude)) %>%
  filter(between(latitude,bb$ymin,bb$ymax),
         between(longitude,bb$xmin,bb$xmax)) %>%
  mutate(E=coalesce(E,1)) %>%
  group_by(Name) %>%
  top_n(1,E) %>%
  expand(latitude,longitude,E,count = seq(1:E)) %>%
  st_as_sf(coords=c("longitude","latitude"),crs=4326,agr="constant") %>%
  ungroup() %>%
  st_jitter(amount=0.005) %>%
  cbind(st_coordinates(.)) %>%
  rename(!!!c("longitude"="X","latitude"="Y"))

  
  #sf::st_as_sf(coords=c("longitude","latitude"),agr="constant",na.fail = FALSE) %>%
g <- ggplot(m) +
  stat_density2d(data=school_data,
                 aes(x=longitude,y=latitude, fill = ..level.., alpha = ..level..),
                 h=c(0.05,0.05*cos(pi/180*49)),#n=1000,
                 bins = 16, geom = "polygon", breaks=c(1,2.5,5,7.5,10,15,20,25,30,40,50,60)) +
  scale_fill_viridis_c(guide=FALSE,option="magma") +
  scale_alpha_continuous(guide=FALSE) +
  geom_roads() +
  geom_water(color="grey",size=0.1) +
  geom_sf(data=school_data,size=0.1,alpha=0.4) +
  geom_sf(fill=NA,color="brown",size=0.1) +
  coord_bbox(metro_van_bbox("tight")) +
  labs(x=NULL,y=NULL,fill=NULL,
       title="Geographic distribution of school exposure notifications",
       caption="MountainMath, Data: BC School Covid Tracker")

g
#r<-graph_to_s3(g,"bccovid","school-tracker-schools.png",width=knitr::opts_chunk$get('fig.width'),height=knitr::opts_chunk$get('fig.height'))

```


## Monthly exposure density
Looking at monthly school exposure denisty gives us some sense of how the distribution of exposure notifications has changed over time. The last month only has partial data and the heat map will appear to have lower values in comparison until the month is complete..


```{r school-tracker-monthly, fig.height=10, fig.height=13}
school_data <- data %>%
  mutate_at(c("latitude","longitude"),as.numeric) %>%
  filter(!is.na(latitude)) %>%
  filter(between(latitude,bb$ymin,bb$ymax),
         between(longitude,bb$xmin,bb$xmax)) %>%
  mutate(Period=strftime(Date,"%b %Y")) %>%
  mutate(Period=factor(Period, levels= arrange(.,Date)$Period %>% unique)) %>%
  st_as_sf(coords=c("longitude","latitude"),crs=4326,agr="constant") %>%
  ungroup() %>%
  st_jitter(amount=0.005) %>%
  cbind(st_coordinates(.)) %>%
  rename(!!!c("longitude"="X","latitude"="Y"))

bb <- c(1,2.5,5,7.5,10,15,20,25,30,40,50,60,100)
bb <- c(1000,2500,5000,7500,10000,15000,20000,25000,30000)
  #sf::st_as_sf(coords=c("longitude","latitude"),agr="constant",na.fail = FALSE) %>%
g <- ggplot(m) +
  stat_density_2d(data=school_data,
                 contour_var="count",
                 aes(x=longitude,y=latitude, fill = ..level.., alpha = ..level..),
                 h=c(0.05,0.05*cos(pi/180*49)),#n=1000,
                 geom = "polygon", breaks=bb
                 ) +
  scale_fill_viridis_c(guide=FALSE,option="magma",limits=c(0,30000)) +
  scale_alpha_continuous(guide=FALSE,limits=c(0,30000)) +
  geom_roads() +
  geom_water(color="grey",size=0.1) +
  geom_sf(data=school_data,size=0.1,alpha=0.4) +
  geom_sf(fill=NA,color="brown",size=0.1) +
  facet_wrap("Period",ncol=2) +
  coord_bbox(metro_van_bbox("tight")) +
  labs(x=NULL,y=NULL,fill=NULL,
       title="Geographic distribution of school exposure notifications",
       caption="MountainMath, Data: BC School Covid Tracker")

g
```


## Children density heat map for comparison

```{r school-tracker-children}
d <- get_census("CA16",regions=list(CMA="59933"),level="DA",geo_format = "sf",vectors=c("v_CA16_25","v_CA16_43","v_CA16_67","v_CA16_70","v_CA16_73"),labels="short", quiet=TRUE) %>%
  mutate(children=select(.,matches("v_")) %>% st_drop_geometry()%>% rowSums(na.rm=TRUE)) 

dd <- d %>% 
  select(children) %>%
  filter(children>0) %>%
  mutate(children=as.integer(children/5)) %>%
  dotdensity::compute_dots("children") %>%
  #st_sample((.)$children) %>%
  st_coordinates() %>%
  as_tibble()
  

g <- ggplot(m) +
  stat_density2d(data=dd,
                 aes(x=X,y=Y, fill = ..level.., alpha = ..level..),
                 h=c(0.05,0.05*cos(pi/180*49)),#n=1000,
                 bins = 16, geom = "polygon") + #, breaks=c(1,2.5,5,7.5,10,12.5,15,17.5,20,25,30,35)) +
  scale_fill_viridis_c(guide=FALSE,option="viridis") +
  scale_alpha_continuous(guide=FALSE) +
  geom_roads() +
  geom_water(color="grey",size=0.1) +
  #geom_sf(data=school_data,size=0.1,alpha=0.4) +
  geom_sf(fill=NA,color="brown",size=0.1) +
  coord_bbox(metro_van_bbox("tight")) +
  labs(x=NULL,y=NULL,fill=NULL,
       title="Geographic distribution of children 5-17yo",
       caption="MountainMath, StatCan Census 2016")

g
#r<-graph_to_s3(g,"bccovid","school-tracker-children.png",width=knitr::opts_chunk$get('fig.width'),height=knitr::opts_chunk$get('fig.height'))
```


