---
title: "BC School Tracker"
author: "Jens von Bergmann"
date: "Last updated at `r format(Sys.time(), '%d %B, %Y - %H:%M',tz='America/Vancouver')`"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	fig.retina = 2,
	dpi = 150,
	fig.width = 7,
	fig.height = 5
)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(rvest)
library(mountainmathHelpers)
library(sf)
library(PROJ)
library(cancensus)
options("cancensus.cache_path"=here::here("data"))

extract_table_data <- function(node){
    h <- node %>% rvest::html_nodes("thead th") %>% rvest::html_text()
  rows <- node %>% rvest::html_nodes("tbody tr")
  data <- rows %>% lapply(function(d) d %>% 
                            rvest::html_nodes("td") %>% 
                            rvest::html_text() %>% 
                            t() %>% 
                            as.data.frame) %>%
    bind_rows() %>%
    setNames(h)
  
  data
}
```

```{r}
get_data_for_page <- function(p=1){
  url = "https://us-east-1-renderer-read.knack.com/v1/scenes/scene_1/views/view_3/records?callback=jQuery17201415845315599671_1606757085608&format=both&rows_per_page=100&sort_field=field_16&sort_order=desc&_=1606757085628"
  url <- paste0(url,"&page=",p)
  
  d <- httr::GET(url,httr::set_cookies( "connect.sid"="s%3A51R3S-8YTv08QV_IejVfqJsW1RxnUdln.WuHEEIrAF6niDEAB3MWgjvWA%2FkzArewbBDZ%2FppCUdVY"),
                 httr::accept("text/javascript, application/javascript, application/ecmascript, application/x-ecmascript, */*; q=0.01"),
                 httr::add_headers("X-Knack-Application-Id"= "5faae3b10442ac00165da195",
                                   "Accept-Encoding"="gzip, deflate, br",
                                   "X-Knack-REST-API-Key"= "renderer",
                                   "x-knack-new-builder"= "true"))
  
  c <- httr::content(d,"text") %>%
    gsub("^.*\\(\\{","{",.) %>%
    gsub("\\}\\);$","}",.) %>%
    jsonlite::fromJSON()
  
  data <- c$records %>% as_tibble()
  
  #print(paste0("Got data for page ",c$current_page," of ",c$total_pages))

  
  attr(data,"total_pages") <- c$total_pages
  data 
}

clean_data <- function(d){
  tibble(Date=as.Date(d$field_16,format="%m/%d/%Y"),
           Name=d$field_13_raw %>% lapply(function(f)f$identifier) %>% as.character,
           `Health Authority`=d$field_26_raw %>% lapply(function(f)f$identifier) %>% as.character,
           Verification=d$field_14_raw %>% lapply(function(f)f$identifier) %>% as.character,
           `Exposure dates`=d$field_15,
           `Exposure count`=d$field_25) %>%
      bind_cols(d$field_19_raw %>% as_tibble) %>%
    mutate(E=str_extract(`Exposure count`,"Exposure \\d+|Secondary \\d+") %>% 
             unlist() %>% 
             gsub("Exposure |Secondary ","",.) %>% 
             as.integer()) %>%
    mutate_at(c("latitude","longitude"),as.numeric)
}

get_all_data <- function(){
  p=1
  raw_data <- get_data_for_page(p)
  data <- clean_data(raw_data)
  while (attr(raw_data,"total_pages")>p) {
    p <- p + 1
    raw_data <- get_data_for_page(p)
    data <- bind_rows(data, clean_data(raw_data))
  }
  data
}

data <- get_all_data()
```

In BC we don't have good data on community level COVID cases. But we can, somewhat cynically, use school exposures as a proxy. The [BC School Covid Tracker project](https://bcschoolcovidtracker.knack.com/bc-school-covid-tracker#home/) has been keeping count of school exposures and meticulously documenting and geocoding the data. That makes is easy for use to create a heatmap of school exposures.

However, this data is not normalized by population, so it should be viewd in context of the school aged (5-17yo) population further down.

## School Exposure Heat Map

```{r school-tracker-schools}
bb<-metro_van_bbox()
m <- cancensus::get_census("CA16",regions=list(CMA="59933"),geo_format = "sf",level="CSD")


school_data <- data %>%
  mutate_at(c("latitude","longitude"),as.numeric) %>%
  filter(!is.na(latitude)) %>%
  filter(between(latitude,bb$ymin,bb$ymax),
         between(longitude,bb$xmin,bb$xmax)) %>%
  mutate(E=coalesce(E,1)) %>%
  group_by(Name) %>%
  top_n(1,E) %>%
  expand(latitude,longitude,E,count = seq(1:E)) %>%
  st_as_sf(coords=c("longitude","latitude"),crs=4326,agr="constant") %>%
  ungroup() %>%
  st_jitter(amount=0.005) %>%
  cbind(st_coordinates(.)) %>%
  rename(!!!c("longitude"="X","latitude"="Y"))

  
  #sf::st_as_sf(coords=c("longitude","latitude"),agr="constant",na.fail = FALSE) %>%
ggplot(m) +
  stat_density2d(data=school_data,
                 aes(x=longitude,y=latitude, fill = ..level.., alpha = ..level..),
                 h=c(0.05,0.05*cos(pi/180*49)),#n=1000,
                 bins = 16, geom = "polygon", breaks=c(1,2.5,5,7.5,10,15,20,25,30,40,50,60)) +
  scale_fill_viridis_c(guide=FALSE,option="magma") +
  scale_alpha_continuous(guide=FALSE) +
  geom_roads() +
  geom_water(color="grey",size=0.1) +
  geom_sf(data=school_data,size=0.1,alpha=0.4) +
  geom_sf(fill=NA,color="brown",size=0.1) +
  coord_bbox(metro_van_bbox("tight")) +
  labs(x=NULL,y=NULL,fill=NULL,
       title="Geographic distribution of school exposure notifications",
       caption="MountainMath, Data: BC School Covid Tracker")
```

## Children density heat map for comparison

```{r school-tracker-children}
d <- get_census("CA16",regions=list(CMA="59933"),level="DA",geo_format = "sf",vectors=c("v_CA16_25","v_CA16_43","v_CA16_67","v_CA16_70","v_CA16_73"),labels="short") %>%
  mutate(children=select(.,matches("v_")) %>% st_drop_geometry()%>% rowSums(na.rm=TRUE)) 

dd <- d %>% 
  select(children) %>%
  filter(children>0) %>%
  mutate(children=as.integer(children/5)) %>%
  dotdensity::compute_dots("children") %>%
  #st_sample((.)$children) %>%
  st_coordinates() %>%
  as_tibble()
  

ggplot(m) +
  stat_density2d(data=dd,
                 aes(x=X,y=Y, fill = ..level.., alpha = ..level..),
                 h=c(0.05,0.05*cos(pi/180*49)),#n=1000,
                 bins = 16, geom = "polygon") + #, breaks=c(1,2.5,5,7.5,10,12.5,15,17.5,20,25,30,35)) +
  scale_fill_viridis_c(guide=FALSE,option="magma") +
  scale_alpha_continuous(guide=FALSE) +
  geom_roads() +
  geom_water(color="grey",size=0.1) +
  #geom_sf(data=school_data,size=0.1,alpha=0.4) +
  geom_sf(fill=NA,color="brown",size=0.1) +
  coord_bbox(metro_van_bbox("tight")) +
  labs(x=NULL,y=NULL,fill=NULL,
       title="Geographic distribution of children 5-17yo",
       caption="MountainMath, StatCan Census 2016")

```


