---
title: "Open Table"
author: "Jens von Bergmann"
date: "Last updated at `r format(Sys.time(), '%d %B, %Y - %H:%M',tz='America/Vancouver')`"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	fig.retina = 2,
	dpi = 150,
	fig.width = 7,
	fig.height = 5
)
library(ggplot2)
library(dplyr)
library(tidyr)
library(CanCovidData)

source(here::here("R/helpers.R"))
```

These graphs show Open Table year over year change in reservations for seated diners and share of restaurants that accept reservations from the [Open Table State of the industry report](https://www.opentable.com/state-of-industry). The code for this notebook is [available for anyone to adapt and use for their own purposes](https://github.com/mountainMath/BCCovidSnippets/blob/main/open_table.Rmd).


## Seated diners from online, phone, and walk-in reservations 

```{r open-table-canada-cities}
open_table <- get_open_table_data("fullbook") %>%
  mutate(value=value/100) %>%
  rename(Name=name)

g <- open_table %>% 
  filter(grepl("Vancouver|Toronto|Calgary|Montr|Edmon|Otta",Name)) %>%
  group_by(Name) %>%
  arrange(Date) %>%
  mutate(Trend=extract_stl_trend(value)) %>%
  ggplot(aes(x=Date,y=Trend,color=Name,group=Name)) +
  geom_point(shape=21,aes(y=value),alpha=0.5) +  
  geom_line(size=0.75) +
  scale_y_continuous(labels=scales::percent)+
  scale_x_date(breaks="months",labels=function(d)strftime(d,"%b")) +
  theme_dark() +
  theme(panel.background = element_rect(fill="#444444"),
        plot.background = element_rect(fill="#444444"),
        legend.background = element_rect(fill="#444444"),
        legend.key = element_rect(fill="#444444"),
        axis.text = element_text(color="whitesmoke"),
        text = element_text(color="whitesmoke"),
        legend.position="bottom") +
  labs(title="OpenTable seated diners restaurant reservations (STL trend lines)",
       x=NULL,y="Year over year change",colour=NULL,caption="MountainMath, Data: OpenTable")

g
#r<-graph_to_s3(g,"bccovid","open_table_canada_cities.png",width=knitr::opts_chunk$get('fig.width'),height=knitr::opts_chunk$get('fig.height'))
```


## Restaurants open for reservations


```{r open-table-canada-cities-2}
open_table <- get_open_table_data("reopening") %>%
  mutate(value=value/100) %>%
  rename(Name=name)

g <- open_table %>% 
  filter(grepl("Vancouver|Toronto|Calgary|Montr|Edmon|Otta",Name)) %>%
  group_by(Name) %>%
  arrange(Date) %>%
  mutate(Trend=extract_stl_trend(value)) %>%
  ggplot(aes(x=Date,y=Trend,color=Name,group=Name)) +
  geom_point(shape=21,aes(y=value),alpha=0.5) +  
  geom_line(size=0.75) +
  scale_y_continuous(labels=scales::percent)+
  scale_x_date(breaks="months",labels=function(d)strftime(d,"%b")) +
  theme_dark() +
  theme(panel.background = element_rect(fill="#444444"),
        plot.background = element_rect(fill="#444444"),
        legend.background = element_rect(fill="#444444"),
        legend.key = element_rect(fill="#444444"),
        axis.text = element_text(color="whitesmoke"),
        text = element_text(color="whitesmoke"),
        legend.position="bottom") +
  labs(title="OpenTable share of restaurants accepting reservations (STL trend lines)",
       x=NULL,y="Share of restaurants accepting reservations",colour=NULL,caption="MountainMath, Data: OpenTable")

g
#r<-graph_to_s3(g,"bccovid","open_table_canada_cities-2.png",width=knitr::opts_chunk$get('fig.width'),height=knitr::opts_chunk$get('fig.height'))
```
