---
title: "Two Covid Canadas"
author: "Jens von Bergmann"
date: "Last updated at `r format(Sys.time(), '%d %B, %Y - %H:%M',tz='America/Vancouver')`"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	fig.retina = 2,
	dpi = 150,
	fig.width = 7,
	fig.height = 5
) 
library(ggplot2)
library(dplyr)
library(tidyr)
library(here)
library(cansim)
library(CanCovidData)

source(here("R/helpers.R"))
```

This notebook has been discontinued 

This notebook shows the confirmed COVID cases for Canadian provinces. The code for this notebook is [available for anyone to adapt and use for their own purposes](https://github.com/mountainMath/BCCovidSnippets/blob/main/two_covid_canadas.Rmd).

```{r}
pop_data <- simpleCache(get_cansim("17-10-0005") %>%
  filter(REF_DATE==2020,`Age group`=="All ages",Sex=="Both sexes") %>%
  select(Province=GEO,Population=VALUE),"prov_pop_data",path = here::here("data"))

covid_data <- get_canada_official_provincial_data() %>%
  mutate(shortProvince=recode(shortProvince,"Nouveau-Brunswick"="NB")) %>%
  mutate(Province=recode(prname,"Nouveau-Brunswick"="New Brunswick")) %>%
  filter(!(Province %in% c("Canada","Repatriated"))) %>%
  mutate(update=coalesce(update,FALSE))


successful_provinces <- c("NL","NT","NS","YT","PE","NB")

province_colours <- c(setNames(RColorBrewer::brewer.pal(6,"Dark2"),c("SK","AB","BC","MB","ON","QC")),
                      setNames(RColorBrewer::brewer.pal(4,"Dark2"),c("NL","NS","PE","NB")))
successful_label <- "Atlantic provinces"

successful_provinces <- c("NT","NS","YT","PE","NB")

province_colours <- c(setNames(RColorBrewer::brewer.pal(7,"Dark2"),c("SK","AB","BC","MB","ON","QC","NS")),
                      setNames(RColorBrewer::brewer.pal(3,"Dark2"),c("NL","PE","NB")))
successful_label <- "Atlantic provinces sans NL"
```

```{r eval=FALSE, include=FALSE}

plot_data <- covid_data %>% 
  select(Date,Province,shortProvince,Cases,update) %>%
  complete(Date,Province) %>%
  mutate(Cases=replace_na(Cases,0)) %>%
  left_join(pop_data,by="Province") %>%
  group_by(Province) %>%
  arrange(desc(Date)) %>%
  filter(cumsum(Cases)>0) %>% # remove trailing zeros
  arrange(Date) %>%
  mutate(Cases=clean_missing_weekend_data(Cases)) %>%
  mutate(incidence=roll::roll_sum(Cases,7)/Population*100000) %>%
  mutate(type=ifelse(shortProvince %in% successful_provinces,"Atlantic bubble & Territories","Rest of Canada"))

plot_data %>%
  filter(Date>=as.Date("2020-03-01"),shortProvince!="CAN") %>%
  #filter(Date>=as.Date("2020-11-01")) %>%
  ggplot(aes(x=Date,y=incidence,group=shortProvince)) +
  geom_line(data=~filter(.,shortProvince %in% successful_provinces),
            color="grey") +
  geom_point(data=~filter(.,shortProvince %in% successful_provinces,Date==max(Date)),
            color="grey") +
  ggrepel::geom_text_repel(data=~filter(.,shortProvince %in% successful_provinces,Date==max(Date)),
            aes(label=shortProvince),nudge_x = 50,min.segment.length = 0,color="grey",direction="y") +
  geom_line(data=~filter(.,!(shortProvince %in% successful_provinces)),
            aes(color=shortProvince)) +
  geom_point(data=~filter(.,!(shortProvince %in% successful_provinces),Date==max(Date)),
            aes(color=shortProvince)) +
  ggrepel::geom_text_repel(data=~filter(.,!(shortProvince %in% successful_provinces),Date==max(Date)),
            aes(color=shortProvince,label=shortProvince),nudge_x = 50,min.segment.length = 0,direction="y") +
  scale_color_brewer(palette = "Dark2",guide=FALSE) +
  facet_wrap("type",ncol=1,scales ="free_y") +
  expand_limits(x=max(plot_data$Date)+7) +
  theme_bw() +
  scale_x_date(breaks="months",labels=function(d)strftime(d,"%b")) +
  labs(title=paste0("7 day incidence for Canadian provinces (as of ",plot_data$Date %>% last,")"),
       x=NULL,y="Cumulative 7 day cases per 100k population",
       color=NULL,
       caption="MountainMath, Data: PHAC")
```

The Atlantic provinces have pursued very different COVID-19 strategies from the other provinces and have seen very different outcomes. The 7-day incidence, that is the cumulative number of cases over the past 7 days per 100,000 population, has been used by many jurisdictions as a key metric to trigger policy interventions.

```{r two-covid-canadas-overview}

plot_data <- covid_data %>% 
  select(Date,Province,shortProvince,Cases,TotalDeaths=Deaths,Deaths=numdeathstoday,update) %>%
  complete(Date,Province) %>%
  mutate(Cases=replace_na(Cases,0),
         Deaths=replace_na(Deaths,0)) %>%
  left_join(pop_data,by="Province") %>%
  group_by(Province) %>%
  arrange(desc(Date)) %>%
  filter(cumsum(update)>0) %>% # remove trailing zeros
  arrange(Date) %>%
  mutate(Cases=clean_missing_weekend_data(Cases)) %>%
  mutate(incidence=roll::roll_sum(Cases,7)/Population*100000) %>%
  mutate(type=ifelse(shortProvince %in% successful_provinces,successful_label,"Other provinces"))

g <- plot_data %>%
  filter(Date>=as.Date("2020-03-01"),shortProvince!="CAN") %>%
  filter(!(shortProvince %in% c("NT","YT","NU"))) %>%
  #filter(Date>=as.Date("2020-11-01")) %>%
  ggplot(aes(x=Date,y=incidence,group=shortProvince,color=type)) +
  geom_line() +
  #geom_point(shape=21) +
  scale_color_manual(values=sanzo::duos$c079 %>% rev) +
  #facet_wrap("type",ncol=1,scales ="free_y") +
  expand_limits(x=max(plot_data$Date)+7) +
  theme_bw() +
  theme(legend.position="bottom") +
  scale_x_date(breaks="months",labels=function(d)strftime(d,"%b")) +
  labs(title="Two COVID Canadas",
       x=NULL,y="Cumulative 7 day cases per 100k population",
       color=NULL,
       caption="MountainMath, Data: PHAC")

g
#r<-graph_to_s3(g,"bccovid","two-covid-canadas.png",width=knitr::opts_chunk$get('fig.width'),height=knitr::opts_chunk$get('fig.height'))
```

For better comparison we can plot the Atlantic provinces and the other provinces on different scales.


```{r two-covid-canadas}
plot_data <- covid_data %>% 
  filter(Province!="Repatriated") %>%
  select(Date,Province,shortProvince,Cases,TotalDeaths=Deaths,Deaths=numdeathstoday,update) %>%
  # complete(Date,Province) %>%
  # mutate(Cases=replace_na(Cases,0),
  #        Deaths=replace_na(Deaths,0)) %>%
  left_join(pop_data,by="Province") %>%
  group_by(Province) %>%
  arrange(desc(Date)) %>%
  filter(cumsum(update)>0) %>% # remove trailing zeros
  arrange(Date) %>%
  mutate(Cases=clean_missing_weekend_data(Cases)) %>%
  mutate(incidence=roll::roll_sum(Cases,7)/Population*100000) %>%
  #mutate(incidence=zoo::rollsum(Cases,7,align="center",fill=as.numeric(NA))/Population*100000) %>%
  mutate(type=ifelse(shortProvince %in% successful_provinces,successful_label,"Other provinces")) 

g <- plot_data %>%
  filter(Date>=as.Date("2020-03-01"),shortProvince!="CAN") %>%
  filter(!(shortProvince %in% c("NT","YT","NU"))) %>%
  #filter(Date>=as.Date("2020-11-01")) %>%
  ggplot(aes(x=Date,y=incidence,group=shortProvince)) +
  geom_line(data=~filter(.,shortProvince %in% successful_provinces),
            aes(color=shortProvince)) +
  geom_point(data=~filter(.,shortProvince %in% successful_provinces,Date==max(Date)),
            aes(color=shortProvince)) +
  ggrepel::geom_text_repel(data=~filter(.,shortProvince %in% successful_provinces,Date==max(Date)),
            aes(label=shortProvince,color=shortProvince),nudge_x = 15,direction="y",
            segment.colour="darkgrey") +
  geom_line(data=~filter(.,!(shortProvince %in% successful_provinces)),
            aes(color=shortProvince)) +
  geom_point(data=~filter(.,!(shortProvince %in% successful_provinces),Date==max(Date)),
            aes(color=shortProvince)) +
  ggrepel::geom_text_repel(data=~filter(.,!(shortProvince %in% successful_provinces),Date==max(Date)),
            aes(color=shortProvince,label=shortProvince),nudge_x = 15,direction="y",
            segment.colour="darkgrey") +
  #scale_color_brewer(palette = "Dark2",guide=FALSE) +
  scale_color_manual(values=province_colours,guide=FALSE) +
  facet_wrap("type",ncol=1,scales ="free_y") +
  expand_limits(x=max(plot_data$Date)+7) +
  theme_bw() +
  scale_x_date(breaks="months",labels=function(d)strftime(d,"%b")) +
  labs(title="Two COVID Canadas",
       x=NULL,y="Cumulative 7 day cases per 100k population",
       color=NULL,
       caption="MountainMath, Data: PHAC")

g
#r<-graph_to_s3(g,"bccovid","two-covid-canadas-overview.png",width=knitr::opts_chunk$get('fig.width'),height=knitr::opts_chunk$get('fig.height'))
```

## Trend lines
Sometimes it is useful to get a clearer view on trend lines. Rolling 7-day sums (like above) or rolling averages (as often emplyed) are a problematic way to represent trend lines as the lag actual trends by 3 days. 

A fairly simple trend line model like a (multiplicative) STL decomposition can extract cleaner trend lines that also cover the most recent 3 days of data, at the expense of a bit of added volatility at the very end of the trend line where the trend line may shift slightly when new data comes in.


```{r two-covid-canadas-trend}
successful_provinces <- c("NL","NT","NS","YT","PE","NB")

province_colours <- c(setNames(RColorBrewer::brewer.pal(6,"Dark2"),c("SK","AB","BC","MB","ON","QC")),
                      setNames(RColorBrewer::brewer.pal(4,"Dark2"),c("NL","NS","PE","NB")))
successful_label <- "Atlantic provinces"

plot_data <- covid_data %>% 
  filter(Province!="Repatriated") %>%
  select(Date,Province,shortProvince,Cases,TotalDeaths=Deaths,Deaths=numdeathstoday,update) %>%
  #complete(Date,Province) %>%
  mutate(Cases=replace_na(Cases,0),
         Deaths=replace_na(Deaths,0)) %>%
  left_join(pop_data,by="Province") %>%
  group_by(Province) %>%
  arrange(desc(Date)) %>%
  filter(cumsum(update)>0) %>% # remove trailing zeros
  arrange(Date) %>%
  mutate(Cases=clean_missing_weekend_data(Cases)) %>%
  mutate(Cases=pmax(0,Cases)) %>%
  mutate(trend=(extract_stl_trend_m(Cases+1)-1)/Population*100000) %>%
  mutate(type=ifelse(shortProvince %in% successful_provinces,successful_label,"Other provinces")) 

g <- plot_data %>%
  filter(Date>=as.Date("2020-03-01"),shortProvince!="CAN") %>%
  filter(!(shortProvince %in% c("NT","YT","NU"))) %>%
  #filter(Date>=as.Date("2020-11-01")) %>%
  ggplot(aes(x=Date,y=trend,group=shortProvince)) +
  geom_line(data=~filter(.,shortProvince %in% successful_provinces),
            aes(color=shortProvince)) +
  geom_point(data=~filter(.,shortProvince %in% successful_provinces,Date==max(Date)),
            aes(color=shortProvince)) +
  ggrepel::geom_text_repel(data=~filter(.,shortProvince %in% successful_provinces,Date==max(Date)),
            aes(label=shortProvince,color=shortProvince),nudge_x = 15,direction="y",
            segment.colour="darkgrey") +
  geom_line(data=~filter(.,!(shortProvince %in% successful_provinces)),
            aes(color=shortProvince)) +
  geom_point(data=~filter(.,!(shortProvince %in% successful_provinces),Date==max(Date)),
            aes(color=shortProvince)) +
  ggrepel::geom_text_repel(data=~filter(.,!(shortProvince %in% successful_provinces),Date==max(Date)),
            aes(color=shortProvince,label=shortProvince),nudge_x = 15,direction="y",
            segment.colour="darkgrey") +
  #scale_color_brewer(palette = "Dark2",guide=FALSE) +
  scale_color_manual(values=province_colours,guide=FALSE) +
  facet_wrap("type",ncol=1,scales ="free_y") +
  expand_limits(x=max(plot_data$Date)+7) +
  theme_bw() +
  scale_x_date(breaks="months",labels=function(d)strftime(d,"%b")) +
  labs(title="Two COVID Canadas (STL trend lines)",
       x=NULL,y="Daily cases per 100k population",
       color=NULL,
       caption="MountainMath, Data: PHAC")

g
#r<-graph_to_s3(g,"bccovid","two-covid-canadas-overview.png",width=knitr::opts_chunk$get('fig.width'),height=knitr::opts_chunk$get('fig.height'))
```
